Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2013-01-14 chi2013-webbox.tex
candb
abstract short

\begin{abstract}
The information processing capabilities of humans enable them to opportunistically draw and integrate knowledge from nearly any information source.  However, the integration of digital, structured data from diverse sources remains difficult, due to problems of heterogeneity that arise when data modelled separately are brought together.  In this paper, we present an investigation of the feasibility of extending the scope of Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, with the objective of allowing  users to take advantage of the vast, emerging ecosystems of structured data currently becoming available.  In this study, we conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources. Observations from these pre-studies led to our designing \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  Our lab study showed, first, that participants readily understood the new interaction mechanisms which were introduced.  Secondly, participants made more carefully justified decisions, even while weighing a greater number of factors, moreover expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up (web page and spreadsheet interface). 
%The information processing capabilities of humans enable people to opportunistically draw and integrate knowledge from nearly any information sources that become available to them.  However, supporting similar, opportunistic integration of digital, structured data from diverse sources remains difficult without tedious, manual integration, due to problems of heterogeneity that arise when data modelled and represented differently are brought together.  In this paper, we investigate the feasibility of minimally extending digital Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, towards enabling users to take advantage of the vast emerging ecosystems of structured data being made available.  We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources. These observations led to our design of \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions during subjective-choice tasks when using DataPalette than when using a control interface, drawing information from more data sources in the process.
% While people naturally draw from diverse information sources in the course of routine decision-making tasks, complementary support for managing heterogeneous data in digital Personal Information Management (PIM) tools remains poor. In this paper, we present an initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web. We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources. These observations led to our design of DataPalette, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions during subjective-choice tasks when using DataPalette than when using a control interface, drawing information from more data sources in the process.

@@

\keywords{Personal Information Management, Personal Data Management, User-driven data integration}

\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}

\terms{Human Factors; Design; Measurement.}

>>>>

\begin{abstract}
The information processing capabilities of humans enable them to opportunistically draw and integrate knowledge from nearly any information source.  However, the integration of digital, structured data from diverse sources remains difficult, due to problems of heterogeneity that arise when data modelled separately are brought together.  In this paper, we present an investigation of the feasibility of extending Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, with the objective of allowing  users to take advantage of the emerging ecosystems of structured data currently becoming available.  In this study, we conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources. Observations from these pre-studies led to \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  Our lab study showed, first, that participants readily understood the new interaction mechanisms which were introduced.  Secondly, participants made more carefully justified decisions, even while weighing a greater number of factors, moreover expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up.

%The information processing capabilities of humans enable people to opportunistically draw and integrate knowledge from nearly any information sources that become available to them.  However, supporting similar, opportunistic integration of digital, structured data from diverse sources remains difficult without tedious, manual integration, due to problems of heterogeneity that arise when data modelled and represented differently are brought together.  In this paper, we investigate the feasibility of minimally extending digital Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, towards enabling users to take advantage of the vast emerging ecosystems of structured data being made available.  We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources. These observations led to our design of \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions during subjective-choice tasks when using DataPalette than when using a control interface, drawing information from more data sources in the process.
% While people naturally draw from diverse information sources in the course of routine decision-making tasks, complementary support for managing heterogeneous data in digital Personal Information Management (PIM) tools remains poor. In this paper, we present an initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web. We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources. These observations led to our design of DataPalette, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions during subjective-choice tasks when using DataPalette than when using a control interface, drawing information from more data sources in the process.

@@

\keywords{Personal information management; end-user data integration; mash-ups; sensemaking with data}

\category{H.5.2.}{Interaction styles}{Direct manipulation}
\category{H.2.5.}{Heterogeneous Databases}{Data translation}
\category{H.5.m.}{Information Interfaces and Presentation}{Miscellaneous}

% \terms{Human Factors; Design; Measurement.}
